### 2020-12-17 EDA_wages_hs2020.Rmd
breaking the spaghetti plot into sample

---
title: "EDA of Wages Tidied Data"
author: "Di Cook, Emi Tanaka, Dewi Lestari Amaliah"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
      keep_tex: yes
      number_sections: yes
      citation_package: biblatex
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(brolgar)
library(visdat)
library(naniar)
library(skimr)
library(kableExtra)
library(patchwork)
library(gghighlight)
library(stickylabeller)
```

# Background

```{r data-load}
load("wages_hs2020.rda")
```

## The original Spaghetty Plot

```{r org_spag, fig.cap = "Wages longitudinal data set"}
fig_all <- ggplot(wages_hs2020) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id))

fig_all_transp <- ggplot(wages_hs2020) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1)

fig_all_lm <- ggplot(wages_hs2020) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  geom_smooth(aes(x = year,
                  y = mean_hourly_wage),
              method = lm,
              se = FALSE)

fig_all + fig_all_transp + fig_all_lm
```
Learn: It is really hard to see individual features since there are a lot observations over time, even if we  change the transparency of the plot.

# Data Description

## Data Structures

```{r dat-structure, fig.cap = "Data structures of wages data"}
visdat::vis_dat(wages_hs2020, 
                warn_large_data = FALSE)
```

## Skim the data

```{r skim}

skim(wages_hs2020) %>%
  as_tibble() %>%
  kable(caption = "The summary of wages data") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

There are 5274 missing values observations in the mean of hourly wage variable or more than 5 percent from the total observations in the data set. This NA is not because of the subject is joining the survey late or out from the survey, but it is the missing value in the series.

# Calculating Features of Data

## Number of Observation per ID

```{r wages_tsibble}
wages <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r}

wages_obs <- wages %>%
  features(mean_hourly_wage, n_obs) %>%
  arrange(n_obs)

ggplot(wages_obs, aes(x = n_obs)) +
  geom_bar() +
  xlab("Number of observations")

```

Learn: 
Different years of observations. The minimum years of observation is 5 years, while the maximum years of observation is 27 years.

```{r first-time, fig.cap="The distribution of sample based on the year to be the first time in the study"}
wages %>%
  features(year, list(min = min)) %>%
  ggplot(aes(x = min)) +
  geom_histogram(binwidth = 1) +
  xlab("First time in study")
```

Learn: Subject starts in the study at different year, most of them start in 1979.

```{r}
wages %>%
  features(year, list(max = max))%>%
  ggplot(aes(x = max)) +
  geom_histogram(binwidth = 1) +
  xlab("Last time in study")
```

Learn: Subject out from the study in the study at different year, most of them are still in the study till the study ends in 2016.

```{r}
wages_year_range <- wages %>%
  features(year, feat_ranges)

ggplot(wages_year_range,
       aes(x = range_diff)) +
  geom_histogram() +
  xlab("Range of year")
```

```{r}
index_regular(wages)
```

Note: the number of observation did not match with the range of year because the year is actually not sequential annually (not regular).

## Minimum, Median, and Maximum of each ID

```{r features}
wages_feature<- wages %>%
  features(mean_hourly_wage, 
           feat_three_num
           )
```

```{r p1}
p1 <- ggplot(wages_feature) +
  geom_density(aes(x = min), fill = "magenta", alpha = 0.5)
```

```{r p2}

wages_feature_long <- wages_feature %>%
  pivot_longer(c(min, med, max), names_to = "feature", values_to = "value")

p2 <- ggplot(wages_feature_long) +
  geom_density(aes(x = value, colour = feature, fill = feature), alpha = 0.3) +
  theme(legend.position = "None")
```

```{r p3}
p3 <- ggplot(transform(wages_feature_long,
                 feature = factor(feature, level = 
                                 c("min",
                                   "med", 
                                   "max")))) +
  geom_line(aes(x = feature,
                y = value, 
                group = id), alpha = 0.5)
```

```{r feature-charts}

p1+p2+p3

```

Learn:

The distribution of min, median, and max of wages is right-skewed, mean that some subjects have an extremely high hourly wages. Further, Some subjects have a minimum wages that is higher than others' maximum wages.

## Is their wage always increase?

```{r}
wages_feat_monotonic <- wages %>%
  features(mean_hourly_wage, 
           feat_monotonic
           )

wages_feat_inc <- wages_feat_monotonic %>%
  filter(increase == TRUE)
```

Learn: Only `r nrow(wages_feat_inc)` subjects out of `r nrow(wages_feat_monotonic)` or `r round(nrow(wages_feat_inc)/nrow(wages_feat_monotonic)*100, 2)` whose mean hourly wage always increase in the period of the study.

## Fit a linear model for each key using `key_slope()`

```{r}
wages_slope <- key_slope(wages, 
                         mean_hourly_wage ~ year) %>%
  left_join(wages, by = "id")

wages_slope
```

```{r decrease-slope, fig.cap= "Subject with negative wages slope"}
ggplot(wages_slope) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  gghighlight(.slope_year < 0)
```

Learn: Some subjects had a negative slope, means that their wages tend to decrease over time.

## Find keys near some summary statistics

```{r}
wages_near <- wages %>%
  key_slope(mean_hourly_wage ~ year) %>%
  keys_near(key = id,
            var = .slope_year)

wages_near
```

# Breaking up the spaghetti

Figure \@ref(fig:org-spag) shows that when the data is presented as a whole spaghetti plot, we can only gain a high level understanding of the data. Hence, we have to break up the spaghetti to see a detailed figure of the individual. We can do this by sampling, stratifying, featuring that strata.

## Sampling

By using `brolgar's` `facet_sample()` function, we can randomly sample the data into a faceted plots.

```{r facet-sample, fig.cap="Mean of hourly wage by sampling the individual"}
ggplot(wages) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  facet_sample()
```
Learn: By default, `facet_sample()` function created 12 facets and 3 sample in each facet. However, as we can see in Figure \@ref(fig:facet-sample), in panel 4 and 10, there are four lines instead of 3. This is due to the missing value in the individual. We can also learn that the subject generally have a positive trend.

## Stratifying

`facet_sample()` only shows part of the data, we can use `facet_strata` to show all the id in different facet.

```{r facet-strata, fig.cap= "Faceted mean hourly wages"}

ggplot(wages) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  facet_strata(n_strata = 100)

```

As we have many ids, it is still hard to see individual data even if it is faceted in to 100 facets. However, we can see that they generally have the same pattern.

## Featuring

We can add the feature in Figure \@ref(fig:facet-strata) by arranging it by year. Here, the data is not randomly allocated into each facet, but they are displayed a specified range of year values.

```{r}
ggplot(wages) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  facet_strata(along = -year)
```
Learn: Most of the samples participated in the study in 1979, this is also inline with the finding in Figure \@ref(fig:first-time). We can see that generally the subjects of the study have the same starting wages but it became vary in the end of the study.