### 2021-01-11 anomaly.Rmd
Calculate the variance of linear model residual.

---
title: "Anomaly_inspection"
author: "Dewi Lestari Amaliah"
date: "09/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brolgar)
```

```{r data}
load("wages_hs2020.rda")
```

```{r wages}
wages <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r lm}
wages_lm <- key_slope(wages, 
                         mean_hourly_wage ~ year) %>%
  left_join(wages, by = "id")
```

```{r}
wages_lm <- wages_lm %>%
  mutate(pred = .intercept + (.slope_year*year))
```

```{r}
wages_res <- wages_lm %>%
  mutate(resid = mean_hourly_wage - pred)
```

```{r}
wages_lm_var <- wages_res %>%
  group_by(id) %>%
  summarise(var = var(mean_hourly_wage,
                      na.rm = TRUE)) %>%
  arrange(desc(var))
```

```{r}
ggplot(filter(wages, id == 8360)) +
  geom_line(aes(x = year,
                y = mean_hourly_wage))
```