### 2021-01-21 code_revisit2.Rmd
Tyding the code and get the raw df of rates and hours

---
title: "code_revisit_2"
author: "Dewi Lestari Amaliah"
date: "21/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(readr)
library(tidyr)
library(tidyverse)
library(brolgar)
```

```{r}

# load the downloaded data 
new_data_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-new_data_qnames.rds"))
categories_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-categories_qnames.rds"))

```

## get hours of work

```{r}

year_A <- c(1979:1987, 1993)
get_hour <- function(year){
  if(year %in% year_A){
   new_data_qnames %>%
    select(CASEID_1979,
            starts_with("QES-52A") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)}
  else{
    new_data_qnames %>%
    select(CASEID_1979,
            starts_with("QES-52D") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)
  }
 }
```

```{r}

hours <- list()

for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016)) {
   hours[[ayear]] <- get_hour(ayear)
}

hours_all <- bind_rows(!!!hours)

```

## get rates

```{r}
get_rate <- function(year) {
  new_data_qnames %>%
     select(CASEID_1979,
            starts_with("HRP") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
    rename(id = CASEID_1979)
}
```

```{r}
rates <- list()

for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016)) {
   rates[[ayear]] <- get_rate(ayear)
}

rates_all <- bind_rows(!!!rates)
```

## join hours and rates

```{r}

hours_wages <- left_join(rates_all, 
                         hours_all, 
                         by = c("id", "year", "job"))

```

## filter how many observations that having rates but "NA" in hours worked

```{r}
na_hours <- hours_wages %>%
  filter(!is.na(rate_per_hour) & is.na(hours_work))

percentage_na_hours <- nrow(na_hours)/ nrow(hours_wages) * 100
```