### 2021-01-29 New_Wages.Rmd
Edit the calculation of wage

---
title: "New_Wages"
author: "Dewi Lestari Amaliah"
date: "29/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(readr)
library(tidyr)
library(tidyverse)
library(brolgar)
library(patchwork)
library(naniar)
```

```{r}

# load the downloaded data 
new_data_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-new_data_qnames.rds"))
categories_qnames <- read_rds(here::here("data-raw/wages-high-school-demo/data-frames/wages-high-school-demo-categories_qnames.rds"))

```

```{r}

year_A <- c(1979:1987, 1993)

get_hour <- function(year){
  if(year %in% year_A){
   new_data_qnames %>%
    select(CASEID_1979,
            starts_with("QES-52A") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)}
  else{
    new_data_qnames %>%
    select(CASEID_1979,
            starts_with("QES-52D") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_", substr(job, 9, 10))) %>%
    rename(id = CASEID_1979)
  }
 }
```

```{r}

hours <- list()

for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016)) {
   hours[[ayear]] <- get_hour(ayear)
}

hours_all <- bind_rows(!!!hours)

```

## get rates

```{r}
get_rate <- function(year) {
  new_data_qnames %>%
     select(CASEID_1979,
            starts_with("HRP") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
    rename(id = CASEID_1979)
}
```

```{r}
rates <- list()

for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016)) {
   rates[[ayear]] <- get_rate(ayear)
}

rates_all <- bind_rows(!!!rates)
```

## join hours and rates

```{r}

hours_wages <- left_join(rates_all, 
                         hours_all, 
                         by = c("id", "year", "job"))
```

```{r}

# calculate number of jobs that a person has in one year
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = length(rate_per_hour))

# filter the observations with available rate per hour
eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  left_join(no_job, by = c("id", "year"))

# calcultae the mean_hourly_wage

# flag1 = code 1 for weighted mean
# code 0 for arithmatic mean

mean_hourly_wage <- 
  eligible_wages %>%
  group_by(id, year) %>%
  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
                        0)) %>%
  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
                        wages))

mean_hourly_wage %>% filter(id == 506)

```
```{r}
visdat::vis_dat(mean_hourly_wage, warn_large_data = FALSE)
```
```{r}
mhw_to_write <- mean_hourly_wage %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages),
            total_hours = sum(hours_work),
            number_of_jobs = mean(no_jobs),
            flag1 = mean(flag1)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  rename(mean_hourly_wage = wages,
         CASEID_1979 = id) %>%
  mutate(flag1 = ifelse(flag1 == 1, "wm",
                        "am"))
```

```{r}
summary(mhw_to_write$mean_hourly_wage)
```

```{r}
ggplot(mhw_to_write) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year)
```
```{r}
#write.csv(mhw_to_write, here::here("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv"))
```