### 2021-01-29 comparing_scenario.Rmd
Comparing the original vs weighted data

---
title: "comparing_scenario"
author: "Dewi Lestari Amaliah"
date: "27/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brolgar)
library(patchwork)
```

```{r}
load("wages_hs2020.rda")
load("wages_hs2020_weighted.rda")
```

```{r}
bp_ori<- ggplot(wages_hs2020) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year) +
  ggtitle("Original")

bp_wei <- ggplot(wages_hs2020_weighted) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year) +
  ggtitle("Weighted")

bp_ori
bp_wei

```

```{r}
wages_ori <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_sc1 <- as_tsibble(x = wages_hs2020_sc1,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_sc2 <- as_tsibble(x = wages_hs2020_sc2,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_weighted <- as_tsibble(x = wages_hs2020_weighted,
                    key = id,
                    index = year,
                    regular = FALSE)
```

```{r}
lp_ori <- ggplot(wages_ori) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Original")

lp_sc1 <- ggplot(wages_sc1) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Scenario 1")

lp_sc2 <- ggplot(wages_sc2) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Scenario 2")

lp_ori + lp_sc1 + lp_sc2

```
```{r}

get_obs <- function(data){
  data %>%
  features(mean_hourly_wage, n_obs) %>%
  arrange(n_obs)
  }

wages_obs_ori <- get_obs(wages_ori)
wages_obs_sc1 <- get_obs(wages_sc1)
wages_obs_sc2 <- get_obs(wages_sc2)
wages_obs_weig <- get_obs(wages_weighted)

nrow(wages_obs_ori)
nrow(wages_obs_sc1)
nrow(wages_obs_sc2)
nrow(wages_obs_weig)
```

```{r}
summary(wages_ori$mean_hourly_wage)
```

```{r}
summary(wages_weighted$mean_hourly_wage)
```