### 2021-02-01 comparing_scenario.Rmd
Comparing original, weighted, and old data

---
title: "comparing_scenario"
author: "Dewi Lestari Amaliah"
date: "27/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brolgar)
library(patchwork)
```

```{r}
load("wages_hs2020.rda")
load("wages_hs2020_weighted.rda")
wages_old <- brolgar::wages %>%
  mutate(mean_hourly_wage = exp(ln_wages))
```

```{r}

# putting into tsibble
wages_ori <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_weighted <- as_tsibble(x = wages_hs2020_weighted,
                    key = id,
                    index = year,
                    regular = FALSE)

wages_old <- as_tsibble(x = wages_old,
                    key = id,
                    index = xp,
                    regular = FALSE)
```

## The summary statistics

```{r}
get_summary <- function(data){
  summary(data$mean_hourly_wage)
}

print("Original data")
get_summary(wages_ori)

print("Weighted data")
get_summary(wages_weighted)

print("Old data")
get_summary(wages_old)

print("Weighted data, extreme value excluded")
get_summary(filter(wages_weighted, flag2 == "non_ext_val"))
```

## Number of ids in each data and number of observations per id

```{r}
get_obs <- function(data){
  data %>%
  features(mean_hourly_wage, n_obs) %>%
  arrange(n_obs)
}

wages_obs_ori <- get_obs(wages_ori)
wages_obs_weig <- get_obs(wages_weighted)
wages_obs_old <- get_obs(wages_old)

nrow(wages_obs_ori)
nrow(wages_obs_weig)
nrow(wages_obs_old)
```

```{r}

get_obs_plot <- function(df, plot_title){
  ggplot(df, aes(x = n_obs)) +
  geom_bar() +
  xlab("Number of observations") +
  ggtitle(plot_title)
}

get_obs_plot(wages_obs_ori, "Original data")
get_obs_plot(wages_obs_weig, "Weighted data")
get_obs_plot(wages_obs_old, "Old data")
```

## First time and last time participated in study (Original vs Weighted data)

```{r}
get_1st_time <- function(df, plot_title){
  df %>%
    features(year, list(min = min)) %>%
    ggplot(aes(x = min)) +
    geom_histogram(binwidth = 1) +
    xlab("First time in study") +
    ggtitle(plot_title)
}

get_1st_time(wages_ori, "Original data")
get_1st_time(wages_weighted, "Weighted data")
```

```{r}
get_last_time <- function(df, plot_title){
  df %>%
    features(year, list(max = max)) %>%
    ggplot(aes(x = max)) +
    geom_histogram(binwidth = 1) +
    xlab("First time in study") +
    ggtitle(plot_title)
}

get_last_time(wages_ori, "Original data")
get_last_time(wages_weighted, "Weighted data")
```

## The spaghetty plot

```{r}

get_sp <- function(df, plot_title, x){
  ggplot(df) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle(plot_title)
}

get_sp(wages_ori, "Original data", year)
get_sp(wages_weighted, "Weighted data", year)

ggplot(wages_old) +
  geom_line(aes(x = xp,
                y = mean_hourly_wage,
                group = id),
            alpha = 0.1) +
  ggtitle("Old data")

```

```{r}
get_sp(filter(wages_weighted, flag2 == "non_ext_val"), "Weighted data - extreme value excluded")
```

## Take a small spoon of those spaghetty

```{r}

set.seed(202102)

plot_sample <- function(df, plot_title){
  ggplot(df) +
  geom_line(aes(x = year,
                y = mean_hourly_wage,
                group = id)) +
  facet_sample() +
  ggtitle(plot_title)
}

plot_sample(wages_ori, "Original data")
plot_sample(wages_weighted, "Weighted data")

ggplot(wages_old) +
  geom_line(aes(x = xp,
                y = mean_hourly_wage,
                group = id)) +
  facet_sample() +
  ggtitle("Old data")

```