### 2021-02-02 School_drop_out.Rmd
Update the code for highschool drop-out.

---
title: "School_drop_out"
author: "Dewi Lestari Amaliah"
date: "13/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(brolgar)
library(patchwork)
library(gghighlight)
```

## filter the observations who are the high school drop out

```{r data}
load("wages_hs2020.rda")

wages <- as_tsibble(x = wages_hs2020,
                    key = id,
                    index = year,
                    regular = FALSE) %>%
  mutate(dob = 1979 - age_1979, 
         age_hgc = yr_hgc - dob)

wages_do <- wages %>%
  filter((hgc %in% c("9TH GRADE", 
                    "10TH GRADE",
                    "11TH GRADE")) |
        (hgc == "12TH GRADE" & 
        age_hgc >= 19))
```

## filter the observations who completed highschool without dropped-out

```{r}
wages_hs_completed <- wages %>%
  filter(hgc == "12TH GRADE" &
           age_hgc < 19)
```

## filter the observations who only completed middle school

```{r}
wages_middle <- wages %>%
  filter(hgc %in% c("6TH GRADE", 
                    "7TH GRADE",
                    "8TH GRADE"))
```

# calculate the summary

```{r}
print("High school droup-out")
summary(wages_do$mean_hourly_wage)
print("Completing high school without being dropped-out")
summary(wages_hs_completed$mean_hourly_wage)
print("completing middle school")
summary(wages_middle$mean_hourly_wage)
```

# USING THE WEIGHTED DATA

## filter the observations who are the high school drop out

```{r data}
load("wages_hs2020_weighted.rda")

wages_w <- as_tsibble(x = wages_hs2020_weighted,
                    key = id,
                    index = year,
                    regular = FALSE) %>%
  mutate(dob = 1979 - age_1979, 
         age_hgc = yr_hgc - dob)

wages_do_w <- wages_w %>%
  filter((hgc %in% c("9TH GRADE", 
                    "10TH GRADE",
                    "11TH GRADE")) |
        (hgc == "12TH GRADE" & 
        age_hgc >= 19))
```

## filter the observations who completed highschool without dropped-out

```{r}
wages_hs_completed_w <- wages_w %>%
  filter(hgc == "12TH GRADE" &
           age_hgc < 19)
```

## filter the observations who only completed middle school

```{r}
wages_middle_w <- wages_w %>%
  filter(hgc %in% c("6TH GRADE", 
                    "7TH GRADE",
                    "8TH GRADE"))
```

# calculate the summary

```{r}
print("High school droup-out")
summary(wages_do_w$mean_hourly_wage)
print("Completing high school without being dropped-out")
summary(wages_hs_completed_w$mean_hourly_wage)
print("completing middle school")
summary(wages_middle_w$mean_hourly_wage)
```