### 2021-02-13 New_Wages.Rmd
Update to survey year 2018 data.

---
title: "New_Wages"
author: "Dewi Lestari Amaliah"
date: "29/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(readr)
library(tidyr)
library(tidyverse)
library(brolgar)
library(patchwork)
library(naniar)
```

## LOAD THE RAW DATA OF DEMOGRAPHIC AND WAGES INFORMATION

```{r}
#source(here::here("data-raw/NLSY79/NLSY79.R"))
```

Creates four files: new_data, has the variables using RNUM;
new_data_qnames, has the data with qnames
categories_vallabels, has RNUM and coding of each variable;
categories_qnames, has qname and coding of each variable;
(qname must be Question name?)

```{r}

# write_rds(as_tibble(new_data_qnames), 
#    here::here("data-raw/NLSY79/NLSY79-new_data_qnames.rds"),       
#       compress = "xz")

# write_rds(as_tibble(categories_qnames), 
#          here::here("data-raw/NLSY79/NLSY79-categories_qnames.rds"), 
#        compress = "xz")

```

```{r}

# load the downloaded data 
new_data_qnames <- read_rds(here::here("data-raw/NLSY79/NLSY79-new_data_qnames.rds"))
categories_qnames <- read_rds(here::here("data-raw/NLSY79/NLSY79-categories_qnames.rds"))

```

## DEMOGRAPHIC DATA PREPROCESSING

```{r}

## tidy the date of birth data

dob_tidy <- new_data_qnames %>%
  select(CASEID_1979,
         starts_with("Q1-3_A~")) %>%
  mutate(dob_year = case_when(
                    # if the years recorded in both sets match, take 79 data
                    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
                    # if the year in the 81 set is missing, take the 79 data
                    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
                    # if the sets don't match for dob year, take the 79 data
                    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
        dob_month = case_when(
                    # if months recorded in both sets match, take 79 data
                    `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
                    # if month in 81 set is missing, take the 79 data
                    is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
                    # if sets don't match for dob month, take the 79 data
                    `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
        # flag if there is a conflict between dob recorded in 79 and 81
        dob_conflict = case_when(     
                      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`) ~ TRUE,
                      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`) ~ TRUE,
                      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) & (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
                      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
  select(CASEID_1979,
         dob_month,
         dob_year,
         dob_conflict)
```

```{r}

## tidy the sex and race data
demog_tidy <- categories_qnames %>%
  select(CASEID_1979,
         SAMPLE_RACE_78SCRN,
         SAMPLE_SEX_1979) %>%
  rename(gender = SAMPLE_SEX_1979,
         race = SAMPLE_RACE_78SCRN)
```

```{r}

# tidy the grade 
demog_education <- new_data_qnames %>%
  as_tibble() %>%
  rename(HGC_2018 = `Q3-4_2018`) %>%
  select(CASEID_1979,
         starts_with("HGCREV"),
         "HGC_2012",
         "HGC_2014",
         "HGC_2016",
         "HGC_2018") %>%
  pivot_longer(!CASEID_1979,
               names_to = "var",
               values_to = "grade") %>%
  separate("var", c("var", "year"), sep = -4) %>%
  filter(!is.na(grade)) %>%
  select(-var)

```

```{r}

## getting the highest year of completed education

highest_year <- demog_education %>%
  group_by(CASEID_1979) %>%
  mutate(hgc_i = max(grade)) %>%
  filter(hgc_i == grade) %>%
  filter(year == first(year)) %>%
  rename(yr_hgc = year) %>%
  select(CASEID_1979, yr_hgc, hgc_i) %>%
  ungroup() %>%
  mutate('hgc' = ifelse(hgc_i == 0, "NONE",
                                  ifelse(hgc_i == 1, "1ST GRADE",
                                         ifelse(hgc_i == 2, "2ND GRADE",
                                                ifelse(hgc_i == 3, "3RD GRADE",
                                                       ifelse(hgc_i >= 4 & hgc_i <= 12, paste0(hgc_i,"TH GRADE"),
                                                              ifelse(hgc_i == 13, "1ST YEAR COL",
                                                                     ifelse(hgc_i == 14, "2ND YEAR COL",
                                                                            ifelse(hgc_i == 15, "3RD YEAR COL",
                                                                                   ifelse(hgc_i == 95, "UNGRADED", paste0((hgc_i - 12), "TH YEAR COL")))))))))))

```

## Join all of the demographic information

```{r}
full_demographics <- full_join(dob_tidy, demog_tidy, by = "CASEID_1979") %>%
  full_join(highest_year, by = "CASEID_1979") %>%
  rename("id" = "CASEID_1979")
```

# WAGES DATA PRE-PROCESSING

## Tidying hour of work data

```{r}

year_A <- c(1979:1987, 1993)

get_hour <- function(year){
  if(year %in% year_A){
   temp <- new_data_qnames %>%
    select(CASEID_1979,
            starts_with("QES-52A") &
              ends_with(as.character(year)))} 
  else{
    temp <- new_data_qnames %>%
    select(CASEID_1979,
            starts_with("QES-52D") &
              ends_with(as.character(year)))} 
    temp %>% 
      pivot_longer(!CASEID_1979,
                 names_to = "job",
                 values_to = "hours_work") %>%
      separate("job", c("job", "year"), sep = -4) %>%
      mutate(job = paste0("job_", substr(job, 9, 10))) %>%
      rename(id = CASEID_1979)

 }
```

```{r}
hours <- list()

for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016, 2018)) {
   hours[[ayear]] <- get_hour(ayear)
}

hours_all <- bind_rows(!!!hours)

```

## Tidying rates per hour data

```{r}
get_rate <- function(year) {
  new_data_qnames %>%
     select(CASEID_1979,
            starts_with("HRP") &
              ends_with(as.character(year))) %>%
    pivot_longer(!CASEID_1979, names_to = "job", values_to = "rate_per_hour") %>%
    separate("job", c("job", "year"), sep = -4) %>%
    mutate(job = paste0("job_0", substr(job, 4, 4))) %>%
    rename(id = CASEID_1979)
}
```

```{r}
rates <- list()

for(ayear in c(1979:1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010, 
               2012, 2014, 2016, 2018)) {
   rates[[ayear]] <- get_rate(ayear)
}

rates_all <- bind_rows(!!!rates)
```

## join hours and rates

```{r}

hours_wages <- left_join(rates_all, 
                         hours_all, 
                         by = c("id", "year", "job")) %>%

# set the 0 value in rate_per_hour as NA 
  mutate(rate_per_hour = ifelse(rate_per_hour == 0, NA,
                                rate_per_hour))
```

```{r}

# calculate number of jobs that a person has in one year
no_job <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  group_by(id, year) %>%
  summarise(no_jobs = length(rate_per_hour))

# filter the observations with available rate per hour
eligible_wages <- hours_wages %>%
  filter(!is.na(rate_per_hour)) %>%
  left_join(no_job, by = c("id", "year"))

# calcultae the mean_hourly_wage

# flag1 = code 1 for weighted mean
# code 0 for arithmatic mean

mean_hourly_wage <- 
  eligible_wages %>%
  group_by(id, year) %>%
  #calculate the weighted mean if the number of jobs > 1
  mutate(wages = ifelse(no_jobs == 1, rate_per_hour/100,
                        weighted.mean(rate_per_hour, hours_work, na.rm = TRUE)/100)) %>%
  #give the flag if it the weighted mean
  mutate(flag1 = ifelse(!is.na(wages) & no_jobs != 1, 1,
                        0)) %>%
  #calculate the arithmetic mean for the na
  mutate(wages = ifelse(is.na(wages), mean(rate_per_hour)/100,
                        wages))

mean_hourly_wage %>% filter(id == 506)

```
```{r}
visdat::vis_dat(mean_hourly_wage, warn_large_data = FALSE)
```
```{r}
mhw_to_write <- mean_hourly_wage %>%
  group_by(id, year) %>%
  summarise(wages = mean(wages),
            total_hours = sum(hours_work),
            number_of_jobs = mean(no_jobs),
            flag1 = mean(flag1)) %>%
  mutate(year = as.numeric(year)) %>%
  ungroup() %>%
  rename(mean_hourly_wage = wages) %>%
  mutate(is_wm = ifelse(flag1 == 1, TRUE,
                        FALSE)) %>%
  select(-flag1)
```

```{r}
summary(mhw_to_write$mean_hourly_wage)
```

```{r}
ggplot(mhw_to_write) +
  geom_boxplot(aes(x = mean_hourly_wage)) +
  facet_wrap(~year)
```
```{r}
write.csv(mhw_to_write, here::here("data-raw/wages-high-school-demo/data-frames/tidy_employment_weighted.csv"))
write.csv(full_demographics, here::here("data-raw/wages-high-school-demo/data-frames/tidy_demographics_v2.csv"))
```