### 2021-02-13 getting_the_data.Rmd

---
title: "Instruction for Getting The Data"
author: "Dewi Lestari Amaliah"
date: "04/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE)
library(tidyverse)
library(brolgar)
```

# Introduction

Data refreshing is the process to update a data set to maintain its timeliness in a consistent and accurate manner. In this case, the data refreshing is performed on the Wages Data from the National Longitudinal Survey of Youth (NLSY) 1979. It is a longitudinal project that follows the lives of a sample of American youth born between 1957-64. The cohort originally included 12,686 respondents ages 14-22 when first interviewed in 1979; after two sub-samples were dropped, 9,964 respondents remain in the eligible samples. Data are now available from Round 1 (1979 survey year) to Round 27 (2016 survey year).

Refreshing the wages data includes the process of cleaning and preprocess the data that downloaded from the NLSY's database to produce the mean hourly wages of the subject along with several demographic variables. The cleaned data set could be used for various purposes, including for teaching and research  purposes, with the demonstration of the usage of `brolgar` package, a package to analyze and visualize longitudinal data with R (Tierney, Cook, and Prvan, 2020). The tidy and cleaned data set is also made to be a package called "yowie".

# Getting The Data

## Getting The Data from The Database

The result of NLYS is stored in a [database](https://www.nlsinfo.org/content/cohorts/nlsy79/get-data) and could be downloaded by variables. In the data base, there are several variables that are ready to be downloaded, such as, education, employment, household, marriage, children, and health. However, we only downloaded these variables:

- Education, Training & Achievement Scores
    -  Education
        -  Summary measures
            - All  schools
                - By year
                    - Highest grade completed (Downloaded all of the 78 variables.)
- Employment
    - Summary measures
        - By job
            - Hours worked (Downloaded all of the 427 variables.)
            - Hourly wages (Downloaded all of the 151 variables.)

            Both hours worked and hourly wages are recorded by the job, up to five jobs for each id/subject. 
- Household, Geography & Contextual Variables
    - Context
        - Summary measures
            - Basic demographics 
              (Downloaded year and month of birth, race, and sex variables.
              There are two versions of the year and month of birth, i.e. 1979 and 1981 data. We downloaded these two versions).

The downloaded data set from the data base are in a csv (wages-high-school-demo.csv) and dat (wages-high-school-demo.dat) format. We will only use the dat format. It also came along with these files:

- wages-high-school-demo.NLSY79: This is the tagset, that can be uploaded to the web site to recreate the data set
- wages-high-school-demo.R: provided by NLSY for reading the data into R and convert the variable's name and its label into something more sensible.

## Data Preprocessing

The downloaded data came with untidy format, where all of the variables for each year being a column (the raw data contains 664 column). We will not use all of the downloaded variables, we will only use the variables as follows:

### Date of Birth

For the Date of Birth data, since there were two versions (1979 and 1981), we only used the 1979 data. Hence, we match both years and flag if it is not match.

```{r dob-tidy}
## tidy the date of birth data
dob_tidy <- new_data_qnames %>%
  select(CASEID_1979,
         starts_with("Q1-3_A~")) %>%
  mutate(dob_year = case_when(
                    # if the years recorded in both sets match, take 79 data
                    `Q1-3_A~Y_1979` == `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`,
                    # if the year in the 81 set is missing, take the 79 data
                    is.na(`Q1-3_A~Y_1981`) ~ `Q1-3_A~Y_1979`,
                    # if the sets don't match for dob year, take the 79 data
                    `Q1-3_A~Y_1979` != `Q1-3_A~Y_1981` ~ `Q1-3_A~Y_1979`),
        dob_month = case_when(
                    # if months recorded in both sets match, take 79 data
                    `Q1-3_A~M_1979` == `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`,
                    # if month in 81 set is missing, take the 79 data
                    is.na(`Q1-3_A~M_1981`) ~ `Q1-3_A~M_1979`,
                    # if sets don't match for dob month, take the 79 data
                    `Q1-3_A~M_1979` != `Q1-3_A~M_1981` ~ `Q1-3_A~M_1979`),
        # flag if there is a conflict between dob recorded in 79 and 81
        dob_conflict = case_when(     
                      (`Q1-3_A~M_1979` != `Q1-3_A~M_1981`) & !is.na(`Q1-3_A~M_1981`) ~ TRUE,
                      (`Q1-3_A~Y_1979` != `Q1-3_A~Y_1981`) & !is.na(`Q1-3_A~Y_1981`) ~ TRUE,
                      (`Q1-3_A~Y_1979` == `Q1-3_A~Y_1981`) & (`Q1-3_A~M_1979` == `Q1-3_A~M_1981`) ~ FALSE,
                      is.na(`Q1-3_A~M_1981`) | is.na(`Q1-3_A~Y_1981`) ~ FALSE)) %>%
  select(CASEID_1979,
         dob_month,
         dob_year,
         dob_conflict)
```

### Sex and Race

For these variables, we only rename the variable to be more clear.

```{r}
## tidy the sex and race data
demog_tidy <- categories_qnames %>%
  select(CASEID_1979,
         SAMPLE_RACE_78SCRN,
         SAMPLE_SEX_1979) %>%
  rename(gender = SAMPLE_SEX_1979,
         race = SAMPLE_RACE_78SCRN)
```

### Education i.e Highest Grade Completed

For the Education i.e. Highest Grade Completed data, we chose the revised May data because the May data seemed to have less missing and presumably the revised data has been checked. However, there was no revised May data for 2012, 2014 and 2016, so we just used the ordinary May data.

Each id will have the highest grade completed for each year. In the final data, we only used the highest grade completed ever and derived the year of when the highest grade completed.

```{r hgc-tidy}
# tidy the grade 
demog_education <- new_data_qnames %>%
  as_tibble() %>%
  select(CASEID_1979,
         starts_with("HGCREV"),
         "HGC_2012",
         "HGC_2014",
         "HGC_2016") %>%
  pivot_longer(!CASEID_1979,
               names_to = "var",
               values_to = "grade") %>%
  separate("var", c("var", "year"), sep = -4) %>%
  filter(!is.na(grade)) %>%
  select(-var)

## getting the highest year of completed education

highest_year <- demog_education %>%
  group_by(CASEID_1979) %>%
  mutate(hgc_i = max(grade)) %>%
  filter(hgc_i == grade) %>%
  filter(year == first(year)) %>%
  rename(yr_hgc = year) %>%
  select(CASEID_1979, yr_hgc, hgc_i) %>%
  ungroup() %>%
  mutate('hgc' = ifelse(hgc_i == 0, "NONE",
                                  ifelse(hgc_i == 1, "1ST GRADE",
                                         ifelse(hgc_i == 2, "2ND GRADE",
                                                ifelse(hgc_i == 3, "3RD GRADE",
                                                       ifelse(hgc_i >= 4 & hgc_i <= 12, paste0(hgc_i,"TH GRADE"),
                                                              ifelse(hgc_i == 13, "1ST YEAR COL",
                                                                     ifelse(hgc_i == 14, "2ND YEAR COL",
                                                                            ifelse(hgc_i == 15, "3RD YEAR COL",
                                                                                   ifelse(hgc_i == 95, "UNGRADED", paste0((hgc_i - 12), "TH YEAR COL")))))))))))
```

### Employment Data (Hours Work and Hourly Wage)

For the Employment i.e. Hours worked per week data, initially only one variable per job, no choice from 1979 to 1987 (QES-52A). From 1988 onward, when we had more options, we chose the variable for total hours including time spent working from home (QES-52D). However, 1993 didn't have all five D variables (the first one and the last one were missing), so we used QES-52A variable instead. Further, 2008 only had jobs 1-4 for the QES-52D variable (whereas the other years had 1-5), so we just used these.

Further, since an id might has more than one job, we calculate the wages for each id in each year using the weighted mean with the hour of work as the weight. However, there are a lot of missing value in hours work data. In this case, if the hourly wage information is available, we will only use the regular/arithmetic mean. Hence, we made a new variable to flag whether the mean hourly wage is a weighted or regular mean. Additionally, if an id only has one job, we will directly use their hourly wages information and recorded as arithmetic mean in the flag.

# Data Dictionary of The Tidy Data

- Kind of the metadata of the tidy data