### 2021-09-19 paper/paper-jsdse/paper-jsdse.Rmd
fixed keywords to YAML, modified JSDSE template to accommodate keywords, fixed abstract, fixed tcolorbox to div fence

---
authors:
  - name: Dewi Amaliah
    thanks: "Test"
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dama0007@student.monash.edu
  - name: Dianne Cook
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dicook@monash.edu
  - name: Emi Tanaka
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: emi.tanaka@monash.edu
  - name: Kate Hyde
    thanks: ""
    department: Department of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: ""
  - name: Nicholas Tierney
    thanks: ""
    department: ""
    affiliation: ""
    location: ""
    email: ""
bibliography: ../references.bib
biblio-style: unsrt
#preamble: >
header-includes:
  - \usepackage{tcolorbox}
output: 
  bookdown::pdf_document2:
    template: TAS-template-for-rmd.tex
    keep_tex: true
date: "`r Sys.Date()`"
always_allow_html: true
---

```{r, include = FALSE}
paper <- here::here("paper/paper.Rmd")
meta <- rmarkdown::yaml_front_matter(paper)
```

---
title: "`r meta$title`"
abstract: >
  `r meta$abstract`
keywords: `r meta$keywords`
---

```{r, child = "../paper.Rmd"}
```