### 2021-09-28 app/app.Rmd
Modify the app.

---
title: "Customize Your Threshold"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(shiny)
library(shinyWidgets)
```

```{r data}
wages_hs <- readRDS(here::here("app/wages_hs.rds")) %>%
    dplyr::select(-`X1.x`, -`X1.y`)
```

Inputs {.sidebar}
-----------------------------------------------------------------------
<br/>

<style>
div.blue { background-color:#FADBD8; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

```{r}
p(icon("info-circle", lib = "font-awesome"), strong("About"))
```

The wages data in `yowie` package uses the weight of robust linear model to detect anomalies and imputes those anomalies with its fitted values. The threshold of weight used is 0.12 to maintain the natural variability of the data. Here, you can customize your own threshold.

</div>

<br/>

```{r}
sliderInput("th", "Threshold:",
            min = 0.01,
            max = 1,
            value = 5)
```

<br/>

Row {data-height=400}
-------------------------------------

### Spaghetti plots

```{r}

thres_data_all <- reactive({
  temp_data <- wages_hs %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))

  wages_compare <- temp_data %>%
    select(id, year, mean_hourly_wage, wages_rlm) %>%
    rename(original = mean_hourly_wage,
           imputed = wages_rlm) %>%
    pivot_longer(c(-id, -year), names_to = "type", values_to = "wages") %>%
    mutate(type = factor(type, levels = c("original", "imputed")))
  })

renderPlot({
  ggplot(thres_data_all()) +
    geom_line(aes(x = year,
                  y = wages,
                  group = id),
              alpha = 0.1) +
    theme_bw() +
    facet_wrap(~type, scales = "free_y") +
    theme(strip.text = element_text(size = 12))
  })
```

### Summary Statistics

```{r}
max_ori <- max(wages_hs$mean_hourly_wage)
min_ori <- round(min(wages_hs$mean_hourly_wage), 2)
med_ori <- round(median(wages_hs$mean_hourly_wage), 2)
mean_ori <- round(mean(wages_hs$mean_hourly_wage), 2)
```

```{r}
pred <- reactive({
  temp_data <- wages_hs %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))
  })
```

<style>
div.blue { background-color:#FADBD8; border-radius: 10px; padding: 10px;}
</style>
<div class = "blue">

```{r}
p(icon("info-circle", lib = "font-awesome"), "Summary statistics of wages in imputed data")
renderText({paste0("Minimum = ", round(min(pred()$wages_rlm), 2), " (Original = 0.01)")})
renderText({paste0("Median = ", round(median(pred()$wages_rlm), 2), " (Original = 7.2)")})
renderText({paste0("Mean = ", round(mean(pred()$wages_rlm), 2), " (Original = 11.87)")})
renderText({paste0("Maximum = ", round(max(pred()$wages_rlm), 2), " (Original = 33400)")})
```

</div>

### Box

```{r}
renderValueBox({
  n_pred <- round(nrow(filter(pred(), is_pred == TRUE))/nrow(pred())*100,2)
  valueBox(
    value = n_pred,
    caption = "percentage of imputed observations",
    icon = "fa-area-chart"
  )
})

```

Row {data-height=650}
-------------------------------------

### Individual plots

```{r}
actionButton("resample", "Sample individuals to display")
```

```{r}

compare_data <- eventReactive(input$resample, {
  sample_id <- sample(unique(wages_hs$id), 16)
  sample <- subset(wages_hs, id %in% sample_id)
})

thres_data <- reactive({
  temp_data <- compare_data() %>%
    mutate(wages_rlm = ifelse(w < input$th  & .fitted >= 0, .fitted,
                                      mean_hourly_wage)) %>%
    mutate(is_pred = ifelse(w < input$th & .fitted >= 0, TRUE, FALSE))

  wages_compare <- temp_data %>%
    select(id, year, mean_hourly_wage, wages_rlm) %>%
    rename(original = mean_hourly_wage,
           imputed = wages_rlm) %>%
    pivot_longer(c(-id, -year), names_to = "type", values_to = "wages") %>%
    mutate(type = factor(type, levels = c("original", "imputed")))
  })

renderPlot({
  ggplot(thres_data()) +
    geom_line(aes(x = year,
                  y = wages,
                  colour = type,
                  linetype = type),
              alpha = 1,
              size = 1) +
    geom_point(aes(x = year,
                   y = wages,
                   colour = type),
               alpha = 0.5,
               size = 1) +
    scale_color_manual(values = c("#229954", "#DC7633")) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 10, size = 5),
          legend.position = "bottom",
          legend.text = element_text(size = 12),
          legend.key.size = unit(6,"line")) +
    facet_wrap(~id, scales = "free_y")
  })

```